---
title: "Practical Machine Learning Course Project - Weight Lifting Exercise"
author: "Juan Carlos González Cardona"
date: "17 de septiembre de 2016"
output: html_document
geometry: margin=1.5cm
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

#Note
All code was putting in Appendix section for improve readability but the response of code was show. Excuse me for my poor english.

#Get data
Get data from repositories and define that separator, decimal and quote character, and NA strings for this case "NA","#DIV/0!","".
```{r getdata, echo=FALSE, warning=FALSE}
urlTraining   <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
urlTesting    <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

wd <- "C:\\Users\\juanc\\Dropbox\\DataScience\\Practical Machine Learning\\Project\\JHU-PracticalMachineLearning"
filesDirectory<- "data"
fileTraining  <- "data\\pml-training.csv"
fileTesting   <- "data\\pml-testing.csv"
setwd(wd)
#rm(list=ls())

if (!file.exists(filesDirectory)) {
  dir.create(filesDirectory)
 }

getData <- function(variableName, url, filename, ...){
  if(!exists(variableName)){
    if (!file.exists(filename)) { 
      download.file(url, filename)
    }
    read.csv2(filename, ...)
  }
}

training <- getData("training", urlTraining, fileTraining, sep = ",", dec=".", quote = "\"", na.strings = c("NA","#DIV/0!",""))
testing <- getData("testing", urlTesting, fileTesting, sep = ",", dec=".", quote = "\"", na.strings = c("NA","#DIV/0!",""))
```


#Preprocess
Only process features, not columns like user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window, num_window.
Overmore, exists a lot of features with a huge NA values, i excluded columns with less then 10% of non NA values.
```{r preprocess, echo=FALSE}
training  <- training[,-c(1:7)]
testing   <- testing[,-c(1:7)]

colNACount    <- colSums(is.na(training)) 
colNoNACount  <- colSums(!is.na(training))

training <- training[, (colNACount / (colNoNACount + colNoNACount)) < 0.10 ]
testing <- testing[, (colNACount / (colNoNACount + colNoNACount)) < 0.10 ]
```

#Split data
Using Caret library to split training set into 2 sets (train 60% and valid 40%) to train models and valid accuracy.
```{r split, echo=FALSE, warning=FALSE}
library(caret)
set.seed(2804) 
inTrain <- createDataPartition(training$classe, p = 0.6, list = FALSE)
train <- training[inTrain, ]
valid <- training[-inTrain, ]
```
Distribution of classe label was balanced thus split dataset was made without special considerations.
```{r plotclass, echo=FALSE, fig.width=8, fig.height=2}
library(ggplot2) 
qplot(classe, data = training, main = "Histogram of Classe for initial training dataset")
qplot(classe, data = train, main = "Histogram of Classe for training dataset")
qplot(classe, data = valid, main = "Histogram of Classe for validate dataset")
```

#Cross Validation
I using K-fold to cross validation to minimize variance. 10 folds take a lot of time for Random Forest and Gradient Boosting. To eval performance of each model i use Accuracy that is the proportion of correct classified observation within all observations,  1 minus accuracy is my expected out of sample error.
To config this k-fold i used trainControl and method "cv".
```{r crossvalidation, echo=FALSE}
trControl <- trainControl(method = "cv", number = 10)
```

#Model Random Forest
Fit to model using Random Forest with default parameters. Accuracy was calculated using confusionMatrix and for this model was 99.17% (95% CI : [0.9895, 0.9936]).
```{r randomforest, echo=FALSE, warning=FALSE}
modRF <- train(classe ~ ., data = train, method = "rf", trControl = trControl)
predRF <- predict(modRF, valid)

confusionMatrix(predRF, valid$classe)
```

#Model Support Vector Machine
Fit to model using Support Vector Machine with default parameters. Accuracy was calculated using confusionMatrix and for this model was 78.1% (95% CI : [0.7717, 0.7901]).
```{r svm, echo=FALSE, warning=FALSE}
modSVM <- train(classe ~ ., data = train, method = "svmLinear", trControl = trControl)
predSVM <- predict(modSVM, valid)
confusionMatrix(predSVM, valid$classe)
```

#Model Stochastic Gradient Boosting
Fit to model using Stochastic Gradient Boosting with default parameters. Accuracy was calculated using confusionMatrix and for this model was 96.1% (95% CI : [0.9565, 0.9652]).
```{r gbm, echo=FALSE, warning=FALSE}
modGBM <- train(classe ~ ., data = train, method = "gbm", trControl = trControl, verbose = FALSE)
predGBM <- predict(modGBM, valid)
confusionMatrix(predGBM, valid$classe)
```

#Decision
The best accuracy was for Random Forest (99.17%), was good for Stochastic Gradient Boosting (96.1%) and disapoint to Support Vector Machine (78.1%).

#Testing
Testing data set was evaluated with Random Forest model and was 100% accuracy when i answered the quiz in Coursera.
```{r testing, echo=FALSE}
print(predict(modRF, newdata=testing))

```

#References
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
http://groupware.les.inf.puc-rio.br/har#ixzz4KeETELOZ



#Appendix
##R code to Get Data
```{r ref.label='getdata', eval=FALSE}

```
##R code to Preprocess
```{r ref.label='preprocess', eval=FALSE}

```
##R code to Split data
```{r ref.label='split', eval=FALSE}

```
```{r ref.label='plotclass', eval=FALSE}

```

##R code to Cross Validation
```{r ref.label='crossvalidation', eval=FALSE}

```
##R code to Model Random Forest
```{r ref.label='randomforest', eval=FALSE}

```
##R code to Model Support Vector Machine
```{r ref.label='svm', eval=FALSE}

```
##R code to Model Stochastic Gradient Boosting
```{r ref.label='gbm', eval=FALSE}

```
##R code to Testing
```{r ref.label='testing', eval=FALSE}

```
